QECO: A QoE-Oriented Computation Offloading Algorithm based on Deep Reinforcement Learning for Mobile Edge Computing

### 2024-10-03 Revision/Paper_Rev.tex
Rev_V4.0
Rev_V5.0

\begin{abstract}
In the realm of mobile edge computing (MEC), efficient computation task offloading plays a pivotal role in ensuring a seamless quality of experience (QoE) for users. Maintaining a high QoE is paramount in today's interconnected world, where users demand reliable services. This challenge stands as one of the most primary key factors contributing to handling dynamic and uncertain mobile environment. In this study, we delve into computation offloading in MEC systems, where strict task processing deadlines and energy constraints can adversely affect the system performance. We formulate the computation task offloading problem as a Markov decision process (MDP) to maximize the long-term QoE of each user individually. We propose a distributed QoE-oriented computation offloading (QECO) algorithm based on deep reinforcement learning (DRL) that empowers mobile devices to make their offloading decisions without requiring knowledge of decisions made by other devices. Through numerical studies, we evaluate the performance of QECO. \textcolor{teal}{Simulation results with 50 mobile devices and 5 edge nodes validate that the QECO algorithm is able to efficiently recognize the mobile devices' demands and exploit the computational resources of edge nodes. Consequently, it can achieve a significant improvement of 37\% -- 75\% in average QoE compared to several benchmark methods and an existing work.}

@@

\begin{table*}[htbp]\textcolor{blue}{
		\renewcommand{\arraystretch}{0.7}
		\captionsetup{name=TABLE}
		\caption{Comparison of Related Works Characteristic}
		\scalebox{1}{%
			\begin{tabular}{ lp{2.2cm}p{3cm}p{3.4cm}p{0.7cm}p{2cm}p{3.2cm}l} 
				\toprule
				\textbf{Paper} & \textbf{Scenario} &  \textbf{Problem} &  \textbf{Objective} &  \textbf{Model} & \textbf{Method} &  \textbf{Drawbacks}  \\ \midrule
				\cite{liao2023online} & Queuing-based MEC & Online computation offloading & Minimize computation delay and energy consumption & MDP & DDPG + DDQN & Only for a singel MEC system model  \\\midrule
				\cite{zhou2021deep} & heterogeneous MEC & Resource allocation and computation offloading & Minimize energy consumption of the entire system & MDP & DDQN & Only for a singel MEC system model \\\midrule				
				\cite{huang2019deep} & Wireless-powered MEC & Computation offloading and resource allocations & Maximize computation rate & MIP & DQN & Not consider delay-sensitive tasks \\\midrule
				\cite{wu2024privacy} & MEC-based IIoT  & Joint power allocation and computation offloading  & Optimization of privacy protection and quality of service. & MDP & Multi-agent actor-critic network& Not take user's demand into consideration \\\midrule
				\cite{zhao2019deep} & MEC & User association and resource allocation & Optimize downlink utility & MDP & Multi-agent D3QN & Not consider system delay and energy consumtion \\\midrule
				\cite{sun2024hierarchical} & MEC & Computation offloading and service caching & Minimize average service delay & MDP & hierarchical DRL & Not consider system energy consumtion \\\midrule
				\cite{wu2023computation} & SDN-based MEC & Stochastic game-based resource allocation & Minimize energy consumption and processing delay & MDP & Multi-agent RL & Not consider tasks with maximum delay tolerance\\\midrule
				\cite{dai2020edge} & MEC & Resource allocation and computation offloading & Minimize energy consumption & MDP & DDPG & Not take user's demand into consideration \\\midrule
				\cite{liu2021learn} & MEC & Resource allocation and computation offloading & Minimize execution cost & POMDP & Multi-agent actor-critic network &  Not consider delay-sensitive tasks \\\midrule
				\cite{gong2022edge} & MEC-based IIoT & Resource allocation and computation offloading & Minimize task computation delay and energy consumption & MDP & DRL & Not take user's demand into consideration\\\midrule
				\cite{gao2022large} ---& Queuing-based heterogeneous MEC & Decentralized computation offloading & Optimize system cost and completion rates & MDP & Multi-agent actor-critic network& Not take user's demand into consideration \\\midrule
				\cite{9253665} &  Queuing-based MEC &  Computation offloading & Minimize drop rate and computation delay. & MDP & D3QN + LSTM & Not consider system energy consumtion\\\midrule
				\cite{wu2024combining} & Queuing-based IIoT MEC & Privacy aware computation offloading & Optimization of computation rate and energy consumption & MDP & Actor-critic network  & Not take user's demand into consideration \\\midrule
				\cite{wu2023multi} & Queuing-based IIoT MEC & Computation offloading & Optimize delay and energy consumption & POMDP & Multi-agent DQN&  Not consider tasks with maximum delay tolerance \\\midrule
				\cite{huang2021deadline} & Queuing-based MEC & Computation offloading & Optimize drop rate and energy consumption & POMDP & DDPG & Not consider computation delay \\ \midrule
				\cite{Bolourian-WCL24} & Queuing-based MEC with EHs& Energy harvesting  computation offloading & Minimize drop rate and energy consumption. & MDP  & DQN & Not consider computation delay  \\\midrule
				QECO &  Queuing-based MEC &  Computation offloading & Maximize the QoE of each MD individually &MDP & D3QN + LSTM & \\
				\toprule
		\end{tabular}}
		\label{table1}}
\end{table*}

@@

\begin{itemize}
	 \item \textit{Task Offloading Problem in the MEC System:} \textcolor{purple}{ We formulate the task offloading problem in queuing-based MEC as an MDP for time-sensitive tasks, taking into account the time-varying system environments (e.g., the arrival of new tasks, and the computational requirement of each task). This approach concentrates on providing high performance in the MEC system while maximizing the long-term QoE, by effectively utilizing the capabilities of resources and handling the dynamic nature of workload at ENs.}

	\item \textit{DRL-based Offloading Algorithm:} To address the problem of long-term QoE maximization, we focus on task completion, task delay, and energy consumption to quantify the MDs' QoE. We propose QECO algorithm based on DRL that empowers each MD to make offloading decisions independently, without prior knowledge of the other MDs' tasks and offloading models. \textcolor{purple}{With a focus on the MD's energy level, our approach leverages dueling double deep Q-network (D3QN) \cite{wang2016dueling} to prioritize and strike an appropriate balance between QoE factors. To handle the uncertainty of workloads at ENs, we integrate long short-term memory (LSTM) \cite{hochreiter1997long} into D3QN, enabling MDs to forecast ENs' workloads over time and improve the estimation of long-term QoE. Additionally, we provide an analysis of the training convergence and computational complexity of the proposed algorithm.}

	\item \textit{Performance Evaluation:} We conduct comprehensive experiments to evaluate the QECO's performance as well as its training convergence under different computation workloads. \textcolor{teal}{The results demonstrate that our algorithm quickly converges compired to DQN and D3QN methods and effectively utilizes the processing capabilities of MDs and ENs, resulting in substantial improvement of at least 37\% in average QoE. This advantage is achieved through a 14\% increase in the number of completed tasks, along with 11.7\% and 10.7\% reductions in task delay and energy consumption, respectively, when compared to the distributed and collective DRL-based offloading algorithm (DCDRL) \cite{qiu2020distributed} and several benchmark methods.}

@@

\begin{figure}
	\captionsetup{name=Fig.}
	\centering
	\includegraphics[width=.9\linewidth]{ queue}
	\vspace*{-1mm}
	\caption{An illustration of MD $i \in \mathcal{I}$ and EN $j \in \mathcal{J}$ in the MEC system.}
	\vspace*{-3mm}
	\label{fig1}
\end{figure}

@@

where $p_i^{\text{C}} = 10^{-27}(f_i)^3$ represents the energy consumption of MD $i$'s CPU frequency []. %\cite{mao2016dynamic}.

@@

In addition to the energy consumed by EN $j$ for task processing, we also take into account the energy consumed by the MD $i$'s user interface in the standby state while waiting for task completion at the EN $j$. We define $E_{i,j}^{\text{I}}(t)$ as the energy consumption associated with the user interface of MD $i \in \mathcal{I}$, which is given by
\begin{alignat}{1}
	E_i^{\text{I}}(t) = D_{i,j}^{\text{E}}(t) p_i^{\text{I}} \tau, %= {\lambda_{i,j}^{\text{E}}(t) \rho_i(t) \over \mathcal{B}_j(t) f_j^{\text{E}}} p_i^{\text{I}},
	\label{16}
\end{alignat}
where $p_i^{\text{I}}$ is the standby energy consumption of MD $i \in \mathcal{I}$.
\begin{alignat}{1}
	E_i^{\text{O}}(t) = E_i^{\text{T}}(t) + \sum_{\mathcal{J}} E_{i,j}^{\text{E}}(t) + E_i^{\text{I}}(t).
	\label{17}
\end{alignat}

@@

$\mathcal{F}_i^t =\bigg \{ t' \bigg|\; t' \leq t,\; \lambda_i(t')>0, \; (1 - x_i(t')) \; l_i^{\text{C}}(t') \; + $ \vspace{-3mm}
\begin{equation}
	\hspace{20mm} x_i(t')\sum\limits_{\mathcal{J}} \sum\limits_{n= t'}^{t}\mathbbm{1}(z_{i,j}^{\text{E}}(n)=z_i(t'))  \; l_{i,j}^{\text{E}}(n) = t \bigg\}.
	\label{27}  
	\nonumber
\end{equation}
Therefore, MD $i$ observes a set of QoEs $\{\boldsymbol{q}_i(t') \mid t' \in \mathcal{F}_i^t\}$ at the beginning of time slot $t+1$, where the set $\mathcal{F}_i^t$ for some $i \in \mathcal{I}$ can be empty. Subsequently, MD $i$ sends its experience $(\boldsymbol{s}_i(t), \boldsymbol{a}_i(t), \boldsymbol{q}_i(t), \boldsymbol{s}_i(t+1))$ to EN $j_i$ for each task $z_i(t')$ in $t' \in \mathcal{F}_i^t$.

@@

\section{Performance Evaluation}\label{section:V}
In this section, we first present the simulation setup and training configuration. \textcolor{teal}{We then illustrate the convergence of the proposed DRL-based QECO algorithm under different optimization methods and hyperparameters, and evaluate its performance in comparison to three baseline schemes in addition to the existing work~\cite{qiu2020distributed}.}

@@

\item \textcolor{teal}{\textbf{DCDRL~\cite{qiu2020distributed}:} This method is a distributed optimization algorithm designed for delay-sensitive tasks in a queuing-based MEC environment, where MDs strategically interact with multiple ENs. We select DCDRL as a benchmark due to its similarity to our work. It is based on the actor-critic framework, which underpins many state-of-the-art DRL algorithms, such as DDPG and Proximal Policy Optimization (PPO)~\cite{30,32}}

@@

We first evaluate the number of completed tasks when comparing our proposed QECO algorithm with the other four schemes. As illustrated in Fig.~\ref{chart1} (a), the QECO algorithm consistently outperforms the benchmark methods when we vary the task arrival rate. At a lower task arrival rate (i.e., 50), most of the methods demonstrate similar proficiency in completing tasks. However, as the task arrival rate increases, the efficiency of QECO becomes more evident. Specifically, when the task arrival rate increases to 250, our algorithm can increase the number of completed tasks by 73\% and 47\% compared to RD and PGOA, respectively.
Similarly, in Fig.~\ref{chart1} (b), as the number of MDs increases, QECO shows significant improvements in the number of completed tasks compared to other methods, especially when faced with a large number of MDs. When there are 110 MDs, our proposed algorithm can effectively increase the number of completed tasks by at least 34\% comparing with other methods. This achievement is attributed to the QECO's ability to effectively handle unknown workloads and prevent congestion at the ENs.

Figs.~\ref{chart2} (a) and ~\ref{chart2} (b) illustrate the overall energy consumption for different values of task arrival rate and the number of MDs, respectively. At the lower task arrival rate, the total energy consumption of all methods is close to each other. The total energy consumption increases when we have a higher task arrival rate.  
%When the task arrival rate increases, the total energy consumption of each method increases differently. It can be found that the production of more computational tasks leads to a continuous increase in energy consumption until the capacity is saturated, and the energy consumption remains constant at its maximum. The QECO algorithm takes into account the battery level of the MD in its decision-making process. 
As can be observed from Fig.~\ref{chart2} (a), at task arrival rate 450, QECO effectively reduces overall energy consumption by 18\% and 15\% compared to RD and PGOA, respectively, as it takes into account the \textcolor{blue}{energy level} of the MD in its decision-making process. However, it consumes more energy compared to LC and FO because they do not utilize all computing resources. In particular, LC only uses the MD's computational resources, while FO utilizes the allocated EN computing resources.

 \begin{figure}[tbp]
	\captionsetup{name=Fig.}
	\begin{minipage}[b]{0.5\linewidth}
		\centering
		\includegraphics[width=\textwidth]{ completed_tasks_2} 
		\textcolor{white}{i}\hspace{0.6cm}(a)
	\end{minipage}
	\hspace{-0.2cm}
	\begin{minipage}[b]{0.5\linewidth}
		\centering
		\includegraphics[width=\textwidth]{ completed_tasks_1}
		\textcolor{white}{i}\hspace{0.6cm}(b)
	\end{minipage}
	\vspace{-0.65cm}
	\caption{\textcolor{blue}{The number of completed tasks under different computation workloads: (a) task arrival rate; (b) the number of MDs.}}
	\label{chart1}
\end{figure}

@@

\begin{figure}[tbp]
	\captionsetup{name=Fig.}
	\begin{minipage}[b]{0.50\linewidth}
		\centering
		\includegraphics[width=\textwidth]{ qoe_1} 		
		\textcolor{white}{i}\hspace{0.6cm}(a)
	\end{minipage}
	\hspace{-0.2cm}
	\begin{minipage}[b]{0.50\linewidth}
		\centering
		\includegraphics[width=\textwidth]{ qoe_2}
		\textcolor{white}{i}\hspace{0.6cm}(b)
	\end{minipage}
	\vspace{-0.7cm}
	\caption{\textcolor{blue}{The average QoE under different computation workloads: (a) task arrival rate; (b) the number of MDs.}}
	\label{chart4}
\end{figure}

In Fig.~\ref{chart2} (b), an increasing trend in overall energy consumption is observed as the number of MDs increases since the number of resources available in the system increases, which leads to higher energy consumption. The QECO algorithm consistently outperforms RD and PGOA methods in overall energy consumption, especially when there are a large number of MDs. Specifically, QECO demonstrates a 27\% and 16\% reduction in overall energy consumption compared to RD and PGOA, respectively, when the number of MDs increases to 110.

As shown in Fig.~\ref{chart3} (a), the QECO algorithm maintains a lower average delay compared to other methods as the task arrival rate increases from 50 to 350. Specifically, when the task arrival rate is 200, it reduces the average delay by at least 12\% compared to other methods. However, for task arrival rates exceeding 350, QECO may experience a higher average delay compared to some of the other methods. This can be attributed to the fact that the other algorithms drop more tasks while our proposed algorithm is capable of completing a higher number of tasks, potentially leading to an increase in average delay. In Fig.~\ref{chart3} (b), as the number of MDs increases, we observe a rising trend in the average delay. It can be inferred that an increase in computational load in the system can lead to higher queuing delays and computations at ENs. Considering the QECO's ability to schedule workloads, when the number of MDs increases from 30 to 110, it consistently maintains a lower average delay which is at least 8\% less than the other methods.

%Fig. 6 illustrates a performance comparison of methods in terms of energy consumption in the entire system, considering an increase in the task arrival rate and the number of MDs. With efficient decision-making, the DRL-based algorithm achieves lower energy consumption in both cases compared to RD and PGOA. However, it consumes more energy compared to LC and FC. This is because LC only utilizes the MD's computational resources, and FC solely utilizes the transmission link. In Fig. 6(a), when the task arrival rate is low, the energy consumption of all methods is close to each other, and as the task arrival rate increases, the energy consumption of each method also increases. It can be observed that the generation of more computational tasks leads to a continuous increase in energy consumption. When the task arrival rate increases from 0.1 to 0.9, the DRL-based algorithm increases by 20\%, while RD and PGOA increase by at least 30\%. Similarly, in Fig. 6(b), the DRL-based algorithm effectively reduces energy consumption, especially when the number of MDs is large.The DRL-based algorithm demonstrates a 15\% reduction in energy consumption compared to RD and a 30\% reduction compared to PGOA as the number of MDs increases to 110.

@@

%The average QoE comparison in Fig.~\ref{chart4} (a) indicates the superiority of the QECO algorithm in providing MDs with a better experience. By taking into account the specific demands of MDs for each key performance metric, it consistently achieves a higher average QoE across various values of arrival task rates when compared to other methods. Specifically, when the task arrival rate is moderate (i.e., 250), QECO increases the average QoE by 57\% and 33\% compared to RD and PGOA, respectively. In Fig.~\ref{chart4} (b), As the number of MDs increases, the potential workload on the ENs also rises, leading to a decrease in the average QoE for each method (excluding LC). It can be observed that an increase in the ENs' workload leads to a reduction in the allocated computing capacity for MDs, resulting in a decrease in the average QoE. However, QECO effectively manages the uncertain load at the ENs. When the number of MDs increases to 90, it increases the average QoE by at least 29\% more than the other methods.

>>>>

\begin{abstract}
In the realm of mobile edge computing (MEC), efficient computation task offloading plays a pivotal role in ensuring a seamless quality of experience (QoE) for users. Maintaining a high QoE is paramount in today's interconnected world, where users demand reliable services. This challenge stands as one of the most primary key factors contributing to handling dynamic and uncertain mobile environment. In this study, we delve into computation offloading in MEC systems, where strict task processing deadlines and energy constraints can adversely affect the system performance. We formulate the computation task offloading problem as a Markov decision process (MDP) to maximize the long-term QoE of each user individually. We propose a distributed QoE-oriented computation offloading (QECO) algorithm based on deep reinforcement learning (DRL) that empowers mobile devices to make their offloading decisions without requiring knowledge of decisions made by other devices. Through numerical studies, we evaluate the performance of QECO. \textcolor{blue}{Simulation results with 50 mobile devices and 5 edge nodes validate that the QECO algorithm is able to efficiently recognize the mobile devices' demands and exploit the computational resources of edge nodes. Consequently, it can achieve a significant improvement of 34.8\% -- 75.6\% in average QoE compared to several benchmark methods and an existing work.}

@@

\begin{table*}[htbp]\textcolor{blue}{
		\renewcommand{\arraystretch}{0.7}
		\captionsetup{name=TABLE}
		\caption{Comparison of Related Works Characteristic}
		\scalebox{1}{%
			\begin{tabular}{ lp{2.2cm}p{3cm}p{3.4cm}p{0.7cm}p{2cm}p{3.2cm}l} 
				\toprule
				\textbf{Paper} & \textbf{Scenario} &  \textbf{Problem} &  \textbf{Objective} &  \textbf{Model} & \textbf{Method} &  \textbf{Drawbacks}  \\ \midrule
				\cite{liao2023online} & Queuing-based MEC & Online computation offloading & Minimize computation delay and energy consumption & MDP & DDPG + DDQN & Only for a singel MEC system model  \\\midrule
				\cite{zhou2021deep} & heterogeneous MEC & Resource allocation and computation offloading & Minimize energy consumption of the entire system & MDP & DDQN & Only for a singel MEC system model \\\midrule				
				\cite{huang2019deep} & Wireless-powered MEC & Computation offloading and resource allocations & Maximize computation rate & MIP & DQN & Not consider delay-sensitive tasks \\\midrule
				\cite{wu2024privacy} & MEC-based IIoT  & Joint power allocation and computation offloading  & Optimization of privacy protection and quality of service. & MDP & Multi-agent actor-critic network& Not take user's demand into consideration \\\midrule
				\cite{zhao2019deep} & MEC & User association and resource allocation & Optimize downlink utility & MDP & Multi-agent D3QN & Not consider system delay and energy consumtion \\\midrule
				\cite{sun2024hierarchical} & MEC & Computation offloading and service caching & Minimize average service delay & MDP & hierarchical DRL & Not consider system energy consumtion \\\midrule
				\cite{wu2023computation} & SDN-based MEC & Stochastic game-based resource allocation & Minimize energy consumption and processing delay & MDP & Multi-agent RL & Not consider tasks with maximum delay tolerance\\\midrule
				\cite{dai2020edge} & MEC & Resource allocation and computation offloading & Minimize energy consumption & MDP & DDPG & Not take user's demand into consideration \\\midrule
				\cite{liu2021learn} & MEC & Resource allocation and computation offloading & Minimize execution cost & POMDP & Multi-agent actor-critic network &  Not consider delay-sensitive tasks \\\midrule
				\cite{gong2022edge} & MEC-based IIoT & Resource allocation and computation offloading & Minimize task computation delay and energy consumption & MDP & DRL & Not take user's demand into consideration\\\midrule
				\cite{gao2022large} & Queuing-based heterogeneous MEC & Decentralized computation offloading & Optimize system cost and completion rates & MDP & Multi-agent actor-critic network& Not take user's demand into consideration \\\midrule
				\cite{9253665} &  Queuing-based MEC &  Computation offloading & Minimize drop rate and computation delay. & MDP & D3QN + LSTM & Not consider system energy consumtion\\\midrule
				\cite{wu2024combining} & Queuing-based IIoT MEC & Privacy aware computation offloading & Optimization of computation rate and energy consumption & MDP & Actor-critic network  & Not take user's demand into consideration \\\midrule
				\cite{wu2023multi} & Queuing-based IIoT MEC & Computation offloading & Optimize delay and energy consumption & POMDP & Multi-agent DQN&  Not consider tasks with maximum delay tolerance \\\midrule
				\cite{huang2021deadline} & Queuing-based MEC & Computation offloading & Optimize drop rate and energy consumption & POMDP & DDPG & Not consider computation delay \\ \midrule
				\cite{Bolourian-WCL24} & Queuing-based MEC with EHs& Energy harvesting  computation offloading & Minimize drop rate and energy consumption. & MDP  & DQN & Not consider computation delay  \\\midrule
				QECO &  Queuing-based MEC &  Computation offloading & Maximize the QoE of each MD individually &MDP & D3QN + LSTM & \\
				\toprule
		\end{tabular}}
		\label{table1}}
\end{table*}

@@

\begin{itemize}
	 \item \textit{Task Offloading Problem in the MEC System:} \textcolor{blue}{ We formulate the task offloading problem in queuing-based MEC as an MDP for time-sensitive tasks, taking into account the time-varying system environments (e.g., the arrival of new tasks, and the computational requirement of each task). This approach concentrates on providing high performance in the MEC system while maximizing the long-term QoE, by effectively utilizing the capabilities of resources and handling the dynamic nature of workload at ENs.}

	\item \textit{DRL-based Offloading Algorithm:} To address the problem of long-term QoE maximization, we focus on task completion, task delay, and energy consumption to quantify the MDs' QoE. We propose QECO algorithm based on DRL that empowers each MD to make offloading decisions independently, without prior knowledge of the other MDs' tasks and offloading models. \textcolor{blue}{With a focus on the MD's energy level, our approach leverages dueling double deep Q-network (D3QN) \cite{wang2016dueling} to prioritize and strike an appropriate balance between QoE factors. To handle the uncertainty of workloads at ENs, we integrate long short-term memory (LSTM) \cite{hochreiter1997long} into D3QN, enabling MDs to forecast ENs' workloads over time and improve the estimation of long-term QoE. Additionally, we provide an analysis of the training convergence and computational complexity of the proposed algorithm.}

\item \textit{Performance Evaluation:} \textcolor{blue}{We conducted comprehensive experiments to evaluate QECO's performance under various computational workloads, as well as its training convergence with different techniques and hyperparameters. The results demonstrate that our algorithm effectively utilizes the processing capabilities of MDs and ENs, addressing the dynamic workloads at ENs. It converges more quickly compared to DQN and D3QN methods, while providing a substantial improvement of at least 34.8\% in average QoE. This advantage is achieved through prioritizing MD's QoE preferences, resulting in improvements of at least 12.2\%, 11.7\%, and 10.8\% in the average number of completed tasks, task delay, and energy consumption, respectively, when compared to the distributed and collective DRL-based offloading algorithm (DCDRL) \cite{qiu2020distributed} and several benchmark methods.}

@@

\begin{figure}
	\captionsetup{name=Fig.}
	\centering
	\includegraphics[width=1\linewidth]{ queue}
	\vspace*{-5mm}
	\caption{An illustration of MD $i \in \mathcal{I}$ and EN $j \in \mathcal{J}$ in the MEC system.}
	\vspace*{-3mm}
	\label{fig1}
\end{figure}

@@

where $p_i^{\text{C}} = 10^{-27}(f_i)^3$ represents the energy consumption of MD $i$'s CPU frequency \cite{mao2016dynamic}.

@@

In addition to the energy consumed by EN $j$ for task processing, we also take into account the energy consumed by the MD $i$'s user interface in the standby state while waiting for task completion at the EN $j$. We define $E_{i,j}^{\text{I}}(t)$ as the energy consumption associated with the user interface of MD $i \in \mathcal{I}$, which is given by
\begin{alignat}{1}
	E_i^{\text{I}}(t) = \textcolor{blue}{\sum_{\mathcal{J}} y_{i,j}(t)} D_{i,j}^{\text{E}}(t) p_i^{\text{I}} \tau, %= {\lambda_{i,j}^{\text{E}}(t) \rho_i(t) \over \mathcal{B}_j(t) f_j^{\text{E}}} p_i^{\text{I}},
	\label{16}
\end{alignat}
where $p_i^{\text{I}}$ is the standby energy consumption of MD $i \in \mathcal{I}$.
\begin{alignat}{1}
	E_i^{\text{O}}(t) = E_i^{\text{T}}(t) + \sum_{\mathcal{J}} \textcolor{blue}{y_{i,j}(t)} E_{i,j}^{\text{E}}(t) + E_i^{\text{I}}(t).
	\label{17}
\end{alignat}

@@

$\mathcal{F}_i^t =\bigg \{ t' \bigg|\; t' \leq t,\; \lambda_i(t')>0, \; (1 - x_i(t')) \; l_i^{\text{C}}(t') \; + $ \vspace{-3mm}
\begin{equation}
	\hspace{10mm} x_i(t')\sum\limits_{\mathcal{J}} \sum\limits_{n= t'}^{t}\mathbbm{1}(z_{i,j}^{\text{E}}(n)=z_i(t'))  \; \textcolor{blue}{y_{i,j}(x)} \, l_{i,j}^{\text{E}}(n) = t \bigg\}.
	\label{27}  
	\nonumber
\end{equation}
Therefore, MD $i$ observes a set of QoEs $\{\boldsymbol{q}_i(t') \mid t' \in \mathcal{F}_i^t\}$ at the beginning of time slot $t+1$, where the set $\mathcal{F}_i^t$ for some $i \in \mathcal{I}$ can be empty. Subsequently, MD $i$ sends its experience $(\boldsymbol{s}_i(t), \boldsymbol{a}_i(t), \boldsymbol{q}_i(t), \boldsymbol{s}_i(t+1))$ to EN $j_i$ for each task $z_i(t')$ in $t' \in \mathcal{F}_i^t$.

@@

\section{Performance Evaluation}\label{section:V}
In this section, we first present the simulation setup and training configuration. \textcolor{blue}{We then illustrate the convergence of the proposed DRL-based QECO algorithm under different optimization methods and hyperparameters, and evaluate its performance in comparison to three baseline schemes in addition to the existing work~\cite{qiu2020distributed}.}

@@

\item \textcolor{blue}{\textbf{DCDRL~\cite{qiu2020distributed}:} This method is a distributed optimization algorithm designed for delay-sensitive tasks in a queuing-based MEC environment, where MDs strategically interact with multiple ENs. We select DCDRL as a benchmark due to its similarity to our work. It is based on the actor-critic framework \cite{konda1999actor}, which underpins many state-of-the-art DRL algorithms, such as DDPG \cite{lillicrap2015continuous}, and Proximal Policy Optimization \cite{schulman2017proximal}}.

@@

 \begin{figure}[tbp]
	\captionsetup{name=Fig.}
	\begin{minipage}[b]{0.5\linewidth}
		\centering
		\includegraphics[width=\textwidth]{ drop_1} 
		\textcolor{white}{i}\hspace{0.6cm}(a)
	\end{minipage}
	\hspace{-0.2cm}
	\begin{minipage}[b]{0.5\linewidth}
		\centering
		\includegraphics[width=\textwidth]{ drop_2}
		\textcolor{white}{i}\hspace{0.6cm}(b)
	\end{minipage}
	\vspace{-0.65cm}
	\caption{\textcolor{blue}{The number of completed tasks under different computation workloads: (a) task arrival rate; (b) the number of MDs.}}
	\label{chart1}
\end{figure}

@@

\begin{figure}[tbp]
	\captionsetup{name=Fig.}
	\begin{minipage}[b]{0.50\linewidth}
		\centering
		\includegraphics[width=\textwidth]{ qoe_2} 		
		\textcolor{white}{i}\hspace{0.6cm}(a)
	\end{minipage}
	\hspace{-0.2cm}
	\begin{minipage}[b]{0.50\linewidth}
		\centering
		\includegraphics[width=\textwidth]{ qoe_1}
		\textcolor{white}{i}\hspace{0.6cm}(b)
	\end{minipage}
	\vspace{-0.7cm}
	\caption{\textcolor{blue}{The average QoE under different computation workloads: (a) task arrival rate; (b) the number of MDs.}}
	\label{chart4}
\end{figure}

We first evaluate the number of completed tasks when comparing our proposed QECO algorithm with the other four schemes. As illustrated in Fig.~\ref{chart1} (a), the QECO algorithm consistently outperforms the benchmark methods when we vary the task arrival rate. At a lower task arrival rate (i.e., 50), most of the methods demonstrate similar proficiency in completing tasks. However, as the task arrival rate increases, the efficiency of QECO becomes more evident. Specifically, when the task arrival rate increases to 250, our algorithm can increase the number of completed tasks by 69.8\% and 29.6\% compared to RD and DCDRL, respectively.
Similarly, in Fig.~\ref{chart1} (b), as the number of MDs increases, QECO shows significant improvements in the number of completed tasks compared to other methods, especially when faced with a large number of MDs. When there are 110 MDs, our proposed algorithm can effectively increase the number of completed tasks by at least 24.8\% comparing with other methods. This achievement is attributed to the QECO's ability to effectively handle unknown workloads and prevent congestion at the ENs.

Figs.~\ref{chart2} (a) and ~\ref{chart2} (b) illustrate the overall energy consumption for different values of task arrival rate and the number of MDs, respectively. At the lower task arrival rate, the total energy consumption of all methods is close to each other. The total energy consumption increases when we have a higher task arrival rate.  
%When the task arrival rate increases, the total energy consumption of each method increases differently. It can be found that the production of more computational tasks leads to a continuous increase in energy consumption until the capacity is saturated, and the energy consumption remains constant at its maximum. The QECO algorithm takes into account the battery level of the MD in its decision-making process. 
As can be observed from Fig.~\ref{chart2} (a), at task arrival rate 450, QECO effectively reduces overall energy consumption by 18.6\% and 16.2\% compared to RD and DCDRL, respectively, as it takes into account the \textcolor{blue}{energy level} of the MD in its decision-making process. However, it consumes more energy compared to LC and FO because they do not utilize all computing resources. In particular, LC only uses the MD's computational resources, while FO utilizes the allocated EN computing resources. -------

In Fig.~\ref{chart2} (b), an increasing trend in overall energy consumption is observed as the number of MDs increases since the number of resources available in the system increases, which leads to higher energy consumption. The QECO algorithm consistently outperforms RD and DCDRL methods in overall energy consumption, especially when there are a large number of MDs. Specifically, QECO demonstrates a 27.4\% and 20.5\% reduction in overall energy consumption compared to RD and DCDRL, respectively, when the number of MDs increases to 110.

As shown in Fig.~\ref{chart3} (a), the QECO algorithm maintains a lower average delay compared to other methods as the task arrival rate increases from 50 to 350. Specifically, when the task arrival rate is 200, it reduces the average delay by at least 12.4\% compared to other methods. However, for task arrival rates exceeding 350, QECO may experience a higher average delay compared to some of the other methods. This can be attributed to the fact that the other algorithms drop more tasks while our proposed algorithm is capable of completing a higher number of tasks, potentially leading to an increase in average delay. In Fig.~\ref{chart3} (b), as the number of MDs increases, we observe a rising trend in the average delay. It can be inferred that an increase in computational load in the system can lead to higher queuing delays and computations at ENs. Considering the QECO's ability to schedule workloads, when the number of MDs increases from 30 to 110, it consistently maintains a lower average delay which is at least 10.8\% less than the other methods.

%Fig. 6 illustrates a performance comparison of methods in terms of energy consumption in the entire system, considering an increase in the task arrival rate and the number of MDs. With efficient decision-making, the DRL-based algorithm achieves lower energy consumption in both cases compared to RD and DCDRL. However, it consumes more energy compared to LC and FC. This is because LC only utilizes the MD's computational resources, and FC solely utilizes the transmission link. In Fig. 6(a), when the task arrival rate is low, the energy consumption of all methods is close to each other, and as the task arrival rate increases, the energy consumption of each method also increases. It can be observed that the generation of more computational tasks leads to a continuous increase in energy consumption. When the task arrival rate increases from 0.1 to 0.9, the DRL-based algorithm increases by 20\%, while RD and DCDRL increase by at least 30\%. Similarly, in Fig. 6(b), the DRL-based algorithm effectively reduces energy consumption, especially when the number of MDs is large.The DRL-based algorithm demonstrates a 15\% reduction in energy consumption compared to RD and a 30\% reduction compared to DCDRL as the number of MDs increases to 110.

@@

%The average QoE comparison in Fig.~\ref{chart4} (a) indicates the superiority of the QECO algorithm in providing MDs with a better experience. By taking into account the specific demands of MDs for each key performance metric, it consistently achieves a higher average QoE across various values of arrival task rates when compared to other methods. Specifically, when the task arrival rate is moderate (i.e., 250), QECO increases the average QoE by 57\% and 33\% compared to RD and DCDRL, respectively. In Fig.~\ref{chart4} (b), As the number of MDs increases, the potential workload on the ENs also rises, leading to a decrease in the average QoE for each method (excluding LC). It can be observed that an increase in the ENs' workload leads to a reduction in the allocated computing capacity for MDs, resulting in a decrease in the average QoE. However, QECO effectively manages the uncertain load at the ENs. When the number of MDs increases to 90, it increases the average QoE by at least 29\% more than the other methods.

We further investigate the overall improvement achieved by the QECO algorithm in comparison to other methods in terms of the average QoE. This metric signifies the advantages MDs obtain by utilizing different algorithms. \textcolor{blue}{Fig.~\ref{chart4} (a) shows the average QoE for different values of the task arrival rate. This figure highlights the superiority of the QECO algorithm in providing MDs with an enhanced experience. At lower task arrival rates (i.e., 50-150), QECO performs optimally and maintains an average QoE of at least 0.72, while the other methods experience a steeper decline, with average QoE dropping to 0--0.54. Specifically, at a task arrival rate of 200, QECO improves the average QoE by 91.7\% compared to DCDRL in the positive average QoE range. As task arrival rates increase to 300, the average QoE significantly decreases for all methods due to increased competition for resources in the MEC system. However, QECO still maintains a positive average QoE, while other methods fall to negative values, with a decline of up to -0.17. For higher task arrival rates (i.e., 350–450), QECO experiences a negative average QoE, but it shows at least a 42.6\% and 24.6\% smaller reduction in average QoE compared to RD and DCDRL, respectively.}

 Fig.~\ref{chart4} (b) illustrates the average QoE when we increase the number of MDs. The EN's workload grows when there are a larger number of MDs, leading to a reduction in the average QoE of all methods except LC. However, QECO effectively manages the uncertain load at the ENs. \textcolor{blue}{When the number of MDs increases from 30 to 110, QECO consistently maintains at least a 29.8\% higher QoE compared to the other methods. Specifically, at a moderate number of 70 MDs, QECO achieves an average QoE of 0.57, showing a 60.3\% improvement compared to DCDRL.} It is worth noting that although improvements in each of the QoE factors can contribute to enhancing system performance, it is essential to consider the user's demands in each time slot. Therefore, the key difference between QECO and other methods is that it prioritizes users' demands, enabling it to strike an appropriate balance among them, ultimately leading to a higher QoE for MDs.
